---
title: "Statistical Inference Part 2:Toothgrowth Data"
author: "Paul Nakonechny"
date: "October 24, 2014"
output: pdf_document
---
This file constitutes the second part of the Statistical Inference Coures Project

####1. Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r}
#Load ToothGrowth data and any relevant packages
data <- ToothGrowth
attach(data)
library(lattice)
library (latticeExtra)
```

```{r}
#Exploratory analysis
#Boxplots

bwplot(dose ~ len, xlab = "Tooth Length", ylab = "Dose", 
       main = "Boxplot")

bwplot(dose ~ len | supp, xlab = "Tooth Length", ylab = "Dose", 
       main = "Tooth length by dose for a given supplement")

#Histogram with density overlay
histogram(len, type = "density", xlab = "Tooth length") + densityplot(len)

qqnorm(len)
qqline(len, probs=c(0.25,0.75))

#We will assume a normal distribution for the purposes of this analysis.
```

####2. Provide a basic summary of the data.

```{r}
#Data summary
#Quantiles
summary(data)

#Variance
datavarsupp <- c(var(len[supp == "OJ"]), var(len[supp == "VC"]))

datavardose <- c(var(len[dose == 0.5]), var(len[dose == 1.0]), var(len[dose == 2.0]))
datavarsupp
datavardose
```
For the analysis comparing differences by supplement, we will not assume equal variances.

For the analysis comparing differences between 0.5 doses and 1.0 doses, we will assume equal variances.

####3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.

```{r}
#Hypothesis tests
#T-test comparing tooth lenth and supplement
tsupp <- t.test(len~supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth)
tsupp

#T-test comparing tooth length under 0.5 and 1.0 doses
tdose.0.5.1 <- t.test(len[dose == 0.5], len[dose == 1.0], paired = FALSE,                                              var.equal = TRUE)
tdose.0.5.1

#T-test comparing tooth length under 0.5 and 2.0 doses
tdose.0.5.2 <- t.test(len[dose == 0.5], len[dose == 2.0], paired = FALSE, 
                      var.equal = FALSE)
tdose.0.5.2 

#T-test comparing tooth length under 1.0 and 2.0 doses
tdose.1.2 <- t.test(len[dose == 1.0], len[dose == 2.0], paired = FALSE, 
                      var.equal = FALSE)
tdose.1.2
```

####4. State your conclusions and the assumptions needed for your conclusions. 
In conclusion, we fail to reject the null hypothesis when comparing tooth length by the type of supplement.

We reject the null hypothesis when comparing tooth length by dosage levels in all cases.

#####Assumptions:
The sample is normally distributed.

For the analysis comparing differences by supplement, I did not assume equal variances.

For the analysis comparing differences between 0.5 doses and 1.0 doses, I assumed equal variances.